---
title: "Breastfeeding Patterns and HMM Replication Code"
author: "Elizabeth Holdsworth"
date: "03-22-2023"
output: html_document
---

# Required Libraries
```{r, warning=FALSE, message=FALSE, error=FALSE}
require("vegan")
require("ANCOMBC")
require("tibble")
require("microbiome")
require("plyr")
require("arsenal")
require("patchwork")
require("tidyverse")
require("psych")
require("corrplot")
require("sjPlot")
require("ggstatsplot")
require("PMCMRplus")
```

# Load Datasets
```{r}
load("genusdata.Rdata")
md <- read.csv("metadata.csv", row.names="ID", as.is=c("X"))
```


# Supplemental Figure 4
```{r, warning=FALSE}
corr <- md[,c("Observed", "Shannon", "shannon_evenness", "zbouts30", "zmomnursew", "zf_allophyscon", "infecorichnessnh", "workoutsidehome", "agemos" )]

cor_test_mat <- corr.test(corr)$p 

M <- cor(corr, use="pairwise.complete.obs")
rownames(M) <- c("Observed richness", "Shannon diversity", "Shannon evenness", "Breastfeeding bouts (z-score)", "Total time breastfeeding (z-score)", "Physical contact frequency (z-score)", "Infant non-household physical ecology", "Maternal work outside home", "Infant age (months)")
colnames(M) <- c("Observed richness", "Shannon diversity", "Shannon evenness", "Breastfeeding bouts (z-score)", "Total time breastfeeding (z-score)", "Physical contact frequency (z-score)", "Infant non-household physical ecology", "Maternal work outside home", "Infant age (months)")

#pdf("corrplot03102023.pdf")
corplot <- corrplot(M, p.mat = cor_test_mat, pch.col = "black", method = "pie", type = "lower", insig = "label_sig", sig.level = c(.001, .01, .05), pch.cex = .9, tl.col = "black",tl.cex = .8)
#dev.off()


```


# Supplemental Figure 1
```{r}
bfplot <- ggplot(md, aes(x=bouts30, y=momnurse60w)) + theme_bw()+ geom_point(alpha=0.75, size=3) + labs(
    x = "Breastfeeding bouts every 30 seconds",
    y = "Total time breastfeeding (minutes)",
    title = "Total time breastfeeding by breastfeeding bouts"
  ) + annotate(geom="text", x=8, y=160, label='r=0.60, p<0.001')

bfplot

#ggsave(bfplot, file="bfplot03102023.png", dpi=600)
```


# Supplemental Tables 2
```{r}
#tested models: in supplemental materials
tmlshannonfb <- lm(Shannon~zbouts30 + agemos + workoutsidehome + zf_allophyscon, data=md)
tmlevennessfb <- lm(shannon_evenness~zbouts30 + agemos + workoutsidehome + zf_allophyscon, data=md)
tmlobservedfb <- lm(Observed~zbouts30 + agemos + workoutsidehome + infecorichnessnh, data=md)
summary(tmlshannonfb)
confint(tmlshannonfb)

summary(tmlevennessfb)
confint(tmlevennessfb)

summary(tmlobservedfb)
confint(tmlobservedfb)

tmlshannonfn <- lm(Shannon~zmomnursew + agemos + workoutsidehome + zf_allophyscon, data=md)
tmlevennessfn <- lm(shannon_evenness~zmomnursew + agemos + workoutsidehome + zf_allophyscon, data=md)
tmlobservedfn <- lm(Observed~zmomnursew + agemos + workoutsidehome + infecorichnessnh, data=md)
summary(tmlshannonfn)
confint(tmlshannonfn)

summary(tmlevennessfn)
confint(tmlevennessfn)

summary(tmlobservedfn)
confint(tmlobservedfn)
```


# Final Multiple Linear Regression Models
```{r}
#Final Models - Used in Manuscript
mlshannonfb <- lm(Shannon~zbouts30 + agemos + workoutsidehome + infecorichnessnh, data=md)
mlevennessfb <- lm(shannon_evenness~zbouts30 + agemos + workoutsidehome + infecorichnessnh, data=md)
mlobservedfb <- lm(Observed~zbouts30 + agemos + workoutsidehome + zf_allophyscon, data=md)
summary(mlshannonfb)
confint(mlshannonfb)

summary(mlevennessfb)
confint(mlevennessfb)

summary(mlobservedfb)
confint(mlobservedfb)

mlshannonfn <- lm(Shannon~zmomnursew + agemos + workoutsidehome + infecorichnessnh, data=md)
mlevennessfn <- lm(shannon_evenness~zmomnursew + agemos + workoutsidehome + infecorichnessnh, data=md)
mlobservedfn <- lm(Observed~zmomnursew + agemos + workoutsidehome + zf_allophyscon, data=md)
summary(mlshannonfn)
confint(mlshannonfn)

summary(mlevennessfn)
confint(mlevennessfn)

summary(mlobservedfn)
confint(mlobservedfn)
```


# Figure 2: Forest plot of regression models
```{r, warning=FALSE}
z<- plot_models(mlobservedfn, mlobservedfb, vline.color = "grey", show.values=TRUE,  show.p = TRUE, p.threshold = c(0.05, 0.01, 0.001), colors="breakfast club", value.size=5, title="(A) Richness", digits=3, axis.labels = c("Breastfeeding bouts (z-score)", "Frequency of non-mother physical contact with infant (z-score)", "Maternal work outside the home", "Infant age (months)", "Total time breastfeeding (z-score)"), show.legend=FALSE) 
z1 <- z + theme_sjplot2(base_size = 20)
z1

y<- plot_models(mlshannonfn, mlshannonfb, vline.color = "grey", show.values=TRUE,  show.p = TRUE, p.threshold = c(0.05, 0.01, 0.001), colors="breakfast club", value.size=5, title="(B) Shannon Diversity", digits=3, axis.labels = c("Breastfeeding bouts (z-score)", "Richness of non-household caregiving network", "Maternal work outside the home", "Infant age (months)", "Total time breastfeeding (z-score)"), show.legend=FALSE)
z2 <- y + theme_sjplot2(base_size = 20) + ylim(-0.5, 0.5)
z2

x<- plot_models(mlevennessfn, mlevennessfb, vline.color = "grey", show.values=TRUE,  show.p = TRUE, p.threshold = c(0.05, 0.01, 0.001), colors="breakfast club", value.size=5, title="(C) Shannon Evenness", digits=3, axis.labels = c("Breastfeeding bouts (z-score)", "Richness of non-household caregiving network", "Maternal work outside the home", "Infant age (months)", "Total time breastfeeding (z-score)"), show.legend=TRUE, legend.title = NULL, m.labels = c("Models including total time breastfeeding", "Models including breastfeeding bouts"))
z3<- x + theme_sjplot2(base_size = 20) + ylim(-0.15, 0.15)
z3

patchwork <- (z1 | z2 | z3)

# Remove axis labels from third subplot
patchwork[[3]] = patchwork[[3]] + theme(axis.text.y = element_blank(),
                                        axis.ticks.y = element_blank(),
                                        axis.title.y = element_blank() )

patchwork

#ggsave(file="Figure1ForestPlots03102023.png", width=19, height=9, units = 'in', dpi=600)

```


# Table 1: Summary table of data
```{r results="asis"}
md$workoutsidehome <- factor(md$workoutsidehome, levels=c(0,1), labels=c("No work outside home", "At least part time work outside home"))
md$zmomnursecutw <- factor(md$zmomnursecutw, levels=c("Low, <-1 z-score", "Medium, [-1,1] z-score", "High, >1 z-score"))
md$zbouts30cut <- factor(md$zbouts30cut, levels=c("Low, <-1 z-score", "Medium, [-1,1] z-score", "High, >1 z-score"))
table1::label(md$bouts30) <- "Breastfeeding Bouts"
table1::label(md$momage2) <- "Maternal age (years)"
table1::label(md$agemos) <- "Infant age (months)"
table1::label(md$f_allophyscon) <- "Frequency of observed physical contact with infant by non-mothers"
table1::label(md$workoutsidehome) <- "Maternal work outside home"
table1::label(md$Shannon) <- "Shannon Diversity"
table1::label(md$Observed) <- "Observed Richness"
table1::label(md$shannon_evenness) <- "Shannon Evenness"
table1::label(md$momnurse60w) <- "Total time breastfeeding"
table1::label(md$infecorichnessnh) <- "Non-household caregiving network"

#across all analytic sample
table1::table1(~bouts30 + momnurse60w + momage2 + agemos + infecorichnessnh + f_allophyscon + workoutsidehome + Shannon + Observed + shannon_evenness, data = md)

#By groups
table_one <- tableby(zbouts30cut ~ bouts30 + momnurse60w + momage2 + agemos + f_allophyscon + kwt(infecorichnessnh) + workoutsidehome + Shannon + Observed + shannon_evenness, data = md) 
summary(table_one, title = "Descriptive Statistics by Breastfeeding Bouts Groups")

table_two <- tableby(zmomnursecutw ~ bouts30 + momnurse60w + momage2 + agemos + f_allophyscon + kwt(infecorichnessnh) + workoutsidehome + Shannon + Observed + shannon_evenness, data = md) 
summary(table_two, title = "Descriptive Statistics by Total Time Breastfeeding Groups")

#for discussion on maternal work outside home
table_three <- tableby(zbouts30cut ~ workoutsidehome, data = md) 
summary(table_three, title = "Maternal work outside home by breastfeeding bouts groups")

table_four <- tableby(zmomnursecutw ~ workoutsidehome, data = md) 
summary(table_four, title = "Maternal work outside home by total time breastfeeding")
```


# ANCOM-BC results for breastfeeding bouts model
```{r}
#reorder groups so medium is reference level
sample_data(genus_ps)$zbouts30cut <- factor(sample_data(genus_ps)$zbouts30cut, levels = c("Medium, [-1,1] z-score", "Low, <-1 z-score", "High, >1 z-score"))
sample_data(genus_ps)$zmomnursecutw <- factor(sample_data(genus_ps)$zmomnursecutw, levels = c("Medium, [-1,1] z-score", "Low, <-1 z-score", "High, >1 z-score"))
colSums(otu_table(genus_ps) != 0)


##ANCOM-BC Model for Bouts Groups
toutz0 = ancombc(phyloseq = genus_ps, formula = "zbouts30cut + agemos + f_allophyscon + workoutsidehome", 
               group = "zbouts30cut", p_adj_method = "fdr", prv_cut = 0.15, lib_cut = 0,
               struc_zero = TRUE, neg_lb = FALSE, tol = 1e-5, 
              max_iter = 100, conserve = TRUE, alpha = 0.05, global = TRUE)

tresz0 = toutz0$res
tresz0
tres_globalz0 = toutz0$res_global
tres_globalz0
tab_coefz0 = tresz0$lfc
tabuntbl <- toutz0$feature_table
```


# ANCOM-BC results for total time breastfeeding model
```{r}
nurseoutz0 = ancombc(phyloseq = genus_ps, formula = "zmomnursecutw + agemos + f_allophyscon + workoutsidehome", 
               group = "zmomnursecutw", p_adj_method = "fdr", prv_cut = 0.15, lib_cut = 0,
               struc_zero = TRUE, neg_lb = FALSE, tol = 1e-5, 
              max_iter = 100, conserve = TRUE, alpha = 0.05, global = TRUE)

nurseresz0 = nurseoutz0$res
nurseresz0
nurseres_globalz0 = nurseoutz0$res_global
nurseres_globalz0
nurseab_coefz0 = nurseresz0$lfc
tabuntblnurse <- nurseoutz0$feature_table
```


# Genera Relative Abundance Information
```{r results="asis", warning=FALSE}
###For tables
#melt into relative abundance by genus
genusrelabundance <- genus_ps %>%
  tax_glom(taxrank = "Genus") %>%                     # agglomerate at genus level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  arrange(Genus) 

#Summary table
table1::table1(~Abundance|Genus, data = genusrelabundance)
```


# Figure 3: ANCOM-BC plot of log-fold changes
```{r}
#Prep ANCOM-BC data for bouts model
sig_taxa = tres_globalz0 %>%
  tibble::rownames_to_column("taxon") %>%
  dplyr::filter(diff_abn == TRUE) %>%
  .$taxon

df_lfc = data.frame(tresz0$lfc[sig_taxa,]) %>% 
  rownames_to_column("taxon_id")
df_lfc$taxon_id = factor(df_lfc$taxon_id, levels = sig_taxa)

df_se = data.frame(tresz0$se[sig_taxa,]) %>% 
  rownames_to_column("taxon_id")


df_se$taxon_id = factor(df_se$taxon_id, levels = sig_taxa)
colnames(df_se)[-1] = paste0(colnames(df_se)[-1], "SE")
df_lfc <- dplyr::rename(df_lfc, lobouts = zbouts30cutLow....1.z.score)
df_lfc <- dplyr::rename(df_lfc, hibouts = zbouts30cutHigh...1.z.score)
df_se <- dplyr::rename(df_se, loboutsse = zbouts30cutLow....1.z.scoreSE)
df_se <- dplyr::rename(df_se, hiboutsse = zbouts30cutHigh...1.z.scoreSE)

lodf_se <- df_se[,c("taxon_id", "loboutsse")]
lodf_se$group <- "Low-Medium"
lodf_se <- dplyr::rename(lodf_se, boutsse = loboutsse)
hidf_se <- df_se[,c("taxon_id", "hiboutsse")]
hidf_se$group <- "High-Medium"
hidf_se <- dplyr::rename(hidf_se, boutsse = hiboutsse)
df_boutsse <- join(hidf_se, lodf_se, type="full")

lodf_lfc <- df_lfc[,c("taxon_id", "lobouts")]
lodf_lfc$group <- "Low-Medium"
lodf_lfc <- dplyr::rename(lodf_lfc, boutslfc = lobouts)
hidf_lfc <- df_lfc[,c("taxon_id", "hibouts")]
hidf_lfc$group <- "High-Medium"
hidf_lfc <- dplyr::rename(hidf_lfc, boutslfc = hibouts)
df_boutslfc <- rbind(lodf_lfc, hidf_lfc)

df_boutslfc$taxon_id <- factor(df_boutslfc$taxon_id, levels = c("Achromobacter", "Acidocella", "Pedobacter", "Micrococcus", "Bifidobacterium"))


p_b = ggplot(data = df_boutslfc, 
           aes(x = boutslfc, y = taxon_id, fill = group)) + 
  geom_bar(stat = "identity", width= 0.67, position = position_dodge(0.7)) + scale_x_continuous(limits=c(-0.45,0.15), breaks=seq(-0.45,0.15,0.1)) +
 geom_text(aes(x=boutslfc/1.25, label = sprintf("%0.2f", round(boutslfc, digits = 2))), position = position_dodge(width = 0.7), color = "black", size = 4) + 
  labs(x = "Log fold change relative to medium group", y = NULL, 
       title = "(B) ANCOM-BC Estimates by Breastfeeding Bouts") + 
  scale_fill_manual(values = c("#18A55E", "#F098D3"), name = element_blank(), guide = guide_legend(reverse = TRUE)) +
  scale_color_discrete(name = NULL) +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1), 
        axis.text.y = element_text(face = "italic"))
p_b

#Prep ANCOM-BC data for total time breastfeeding
nsig_taxa = nurseres_globalz0 %>%
  tibble::rownames_to_column("taxon") %>%
  dplyr::filter(diff_abn == TRUE) %>%
  .$taxon

ndf_lfc = data.frame(nurseresz0$lfc[nsig_taxa,]) %>% 
  rownames_to_column("taxon_id")
ndf_lfc$taxon_id = factor(ndf_lfc$taxon_id, levels = nsig_taxa)

ndf_se = data.frame(nurseresz0$se[nsig_taxa,]) %>% 
  rownames_to_column("taxon_id")

ndf_se$taxon_id = factor(ndf_se$taxon_id, levels = nsig_taxa)
colnames(ndf_se)[-1] = paste0(colnames(ndf_se)[-1], "SE")
ndf_lfc <- dplyr::rename(ndf_lfc, lonurse = zmomnursecutwLow....1.z.score)
ndf_lfc <- dplyr::rename(ndf_lfc, hinurse = zmomnursecutwHigh...1.z.score)
ndf_se <- dplyr::rename(ndf_se, lonursese = zmomnursecutwLow....1.z.scoreSE)
ndf_se <- dplyr::rename(ndf_se, hinursese = zmomnursecutwHigh...1.z.scoreSE)

londf_se <- ndf_se[,c("taxon_id", "lonursese")]
londf_se$group <- "Low-Medium"
londf_se <- dplyr::rename(londf_se, nursese = lonursese)
hindf_se <- ndf_se[,c("taxon_id", "hinursese")]
hindf_se$group <- "High-Medium"
hindf_se <- dplyr::rename(hindf_se, nursese = hinursese)
df_nursese <- join(hindf_se, londf_se, type="full")

londf_lfc <- ndf_lfc[,c("taxon_id", "lonurse")]
londf_lfc$group <- "Low-Medium"
londf_lfc <- dplyr::rename(londf_lfc, nurselfc = lonurse)
hindf_lfc <- ndf_lfc[,c("taxon_id", "hinurse")]
hindf_lfc$group <- "High-Medium"
hindf_lfc <- dplyr::rename(hindf_lfc, nurselfc = hinurse)
df_nurselfc <- rbind(londf_lfc, hindf_lfc)

df_nurselfc$taxon_id <- factor(df_nurselfc$taxon_id, levels = c("Stenotrophomonas", "Rugamonas", "Pedobacter", "Agreia", "Bifidobacterium"))

#total time breastfeeding ancom-bc plot
np_b = ggplot(data = df_nurselfc, 
           aes(x = nurselfc, y = taxon_id, fill = group)) + 
  geom_bar(stat = "identity", width= 0.67, position = position_dodge(0.7)) + scale_x_continuous(limits=c(-1,0), breaks=seq(-1,0,0.1)) +
  #geom_errorbarh(aes(xmin = nurselfc - nursese, xmax = nurselfc + nursese), width = 0.2,
  #              position = position_dodge(0.8), color = "black") + 
  geom_text(aes(x=nurselfc/1.25, label = sprintf("%0.2f", round(nurselfc, digits = 2))), position = position_dodge(width = 0.7), color = "black", size = 4) + 
  labs(x = "Log fold change relative to medium group", y = NULL, 
       title = "(A) ANCOM-BC Estimates by Total Time Breastfeeding") + 
  scale_fill_manual(values = c("#18A55E", "#F098D3"), name = element_blank(), guide = guide_legend(reverse = TRUE)) +
  scale_color_discrete(name = NULL) +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1), 
        axis.text.y = element_text(face = "italic"), 
        legend.position = "none")
np_b


##Put together
logpatch <- np_b + p_b

logpatch

#ggsave(logpatch, file="Log fig 03032023.png", height = 5, width = 13, dpi=500)
```


# Figure 1: Bar chart of genera
```{r warning=FALSE}
#settings for plots
brks <- c(0, .20, .40, .60, .80, 1)
y_labels <- c("0", "20", "40", "60", "80", "100")
customcol <- c("#c5723e","#ff98e3","#78cf63","#9675ff","#ffd431","#740067","#b4b15b","#ff3126","#00bcc6","#670100","#708dd4","#0e4b00","#c74f71","#0026b7")

custom_col95 <- c(
 "#6531b9", "#00B0F0", "#FFFF00", "#96D050", "#CC3399", "#375623", "#FFC000", "#0070C0", "#990033", "#00B050",
 "#FF00FF", "#66FF99", "#F96E05", "#FFFF99", "#000000", "#0000FF", "#FF7C80", "#CC66FF","#00FF00", "#002060",
"#5F5F5F", "#FF0066", "#666633", "#FF99FF", "#CCCC00", "#66FFFF", "#660033","#D9D9D9", "#666699", "#660066",
"#781156", "#6CABEA", "#3FE491", "#784511", "#EA6C81", "#A51876","#98C4F0", "#6CEAAB", "#A55E18", "#F098A7",
"#D21E96", "#117878" ,"#98F0C4", "#D2781E", "#E43FAD","#18A5A5", "#787811", "#E4913F", "#EA6CC0", "#3FE4E4",
"#A5A518", "#EAAB6C", "#F098D3",  "#114578","#185EA5", "#1E78D2", "#3F91E4", "#6CEAEA", "#98F0F0", "#117845",
"#18A55E", "#1ED278","#D2D21E","#E4E43F","#EAEA6C", "#F0F098","#F7F7C5", "#F0C498", "#781122", "#A5182F",
"#D21E2C", "#E43F5B",
"#FF83FA","#9FB6CD","#EEE685","#EE5C42","#71C671","#FF4040","#CDC5BF","#C0FF3E","#308014","#00EEEE",
"#9B30FF","#CDC1C5","#4876FF" ,"#EEEE00","#FF7F00","#B3EE3A",
"#F0C498","#781122","#A5182F","#D21E2C",
"#E43F5B", "#A5A518","#EAAB6C", "#F098D3"
)

###Prep data for plots
#Identify taxa to label <5% or 1% across total sample
spectaxa<-aggregate(Abundance~Genus, mean, data=genusrelabundance)
taxa5 <- spectaxa$Genus[spectaxa$Abundance < 0.05]
taxa1 <- spectaxa$Genus[spectaxa$Abundance < 0.01]

#order genera by abundance for use in sample plot
spectaxa$Genus[spectaxa$Abundance < 0.01] <- "< 1% abund."
spectaxa$genuso <- factor(spectaxa$Genus, levels=unique(spectaxa$Genus[order(spectaxa$Abundance)]), ordered=TRUE)

#create dataset of taxa by sample
taxameanssample<-aggregate(Abundance~Genus + Sample, mean, data=genusrelabundance)
#find sample order by Staphylococcus abundance before assigning the <1% abund label, so the x-axis is ordered by Staphylococcus abundance
taxameanssample$Sample <- as.factor(taxameanssample$Sample)
test<- taxameanssample[taxameanssample$Genus=="Staphylococcus",]
test$fn <- factor(test$Sample, levels=unique(test$Sample[order(test$Abundance)]), ordered=TRUE)
test$fn <- factor(test$fn, levels=rev(levels(test$fn)))
sort(test$fn, decreasing=TRUE)

#assign <1% abundance to those taxa and order the x-axis and genera
taxameanssample$Genus[taxameanssample$Genus %in% taxa1 == TRUE] <- "< 1% abund."
taxameanssample$percent <- taxameanssample$Abundance*100
taxa_list_truncsample <- taxameanssample %>% group_by("Sample")
taxa_list_truncsample$Sample <- ordered(taxa_list_truncsample$Sample, levels = levels(test$fn))
taxa_list_truncsample$Genus <- ordered(taxa_list_truncsample$Genus, levels = levels(spectaxa$genuso))

##Figure 2: Sample plot
genussample <- ggplot(taxa_list_truncsample, aes(x=Sample, y = Abundance, fill=factor(Genus), color=customcol)) + 
  geom_bar(stat = "identity", position = position_fill(reverse = FALSE), color="NA") +
  scale_fill_manual(values = customcol) +
  ggtitle("Genera Relative Abundance") + labs(y="Relative abundance (%)", x="Sample") + guides(fill=guide_legend(title="Genera", ncol = 1, reverse = FALSE)) + theme(axis.title.x = element_text(vjust=-1.2), legend.text = element_text(face="italic"), panel.background = element_blank(), axis.line = element_line(colour = "grey"), legend.key.size = unit(0.4, "cm")) + scale_y_continuous(breaks=brks, labels = y_labels) + theme(text = element_text(size = 20), axis.text.x = element_blank()) 

#ggsave("stackedbarchartsample03102023.png", plot=genussample, height = 7, width = 7, unit="in", dpi=500)

genussample
```

# Supplemental Figure 5
```{r}
##taxa by total time breastfeeding groups, for supplemental materials
taxameans<-aggregate(Abundance~zmomnursecutw+Genus, mean, data=genusrelabundance)
taxameans$Genus[taxameans$Genus %in% taxa1 == TRUE] <- "< 1% abund."
taxameans$percent <- taxameans$Abundance*100
taxa_list_trunc1 <- taxameans %>% group_by("zmomnursecutw")
taxa_list_trunc1$zmomnursecutw <- factor(taxa_list_trunc1$zmomnursecutw, levels = c("Low, <-1 z-score","Medium, [-1,1] z-score","High, >1 z-score"))

bfbarchart <- ggplot(taxa_list_trunc1, aes(x=zmomnursecutw, y = Abundance, fill=factor(Genus), color=custom_col95)) + 
  geom_bar(stat = "identity", position = position_fill(reverse = FALSE), color="NA") +
  scale_fill_manual(values = custom_col95) +
  ggtitle("Genera Relative Abundance") + labs(y="Relative abundance (%)", x="Total Time Breastfeeding Z-Score Groups") + guides(fill=guide_legend(title="Genera", ncol = 1, reverse = FALSE)) + theme(axis.title.x = element_text(vjust=-1.2), legend.text = element_text(face="italic"), panel.background = element_blank(), axis.line = element_line(colour = "grey"), legend.key.size = unit(0.4, "cm")) + scale_y_continuous(breaks=brks, labels = y_labels) + theme(text = element_text(size = 20)) 

#ggsave("stackedbarchartnurse03102023.png", plot=bfbarchart, height = 8, width = 11, unit="in", dpi=500)

bfbarchart

##Taxa by bf bouts groups, for supplemental materials
taxameansbf<-aggregate(Abundance~zbouts30cut+Genus, mean, data=genusrelabundance)
taxameansbf$Genus[taxameansbf$Genus %in% taxa1 == TRUE] <- "< 1% abund."
taxameansbf$percent <- taxameansbf$Abundance*100
taxa_list_truncbf <- taxameansbf %>% group_by("zbouts30cut")
taxa_list_truncbf$zbouts30cut <- factor(taxa_list_truncbf$zbouts30cut, levels = c("Low, <-1 z-score","Medium, [-1,1] z-score","High, >1 z-score"))

genusbouts <- ggplot(taxa_list_truncbf, aes(x=zbouts30cut, y = Abundance, fill=factor(Genus), color=custom_col95)) + 
  geom_bar(stat = "identity", position = position_fill(reverse = FALSE), color="NA") +
  scale_fill_manual(values = custom_col95) +
  ggtitle("Genera Relative Abundance") + labs(y="Relative abundance (%)", x="Breastfeeding Bouts Z-Score Groups") + guides(fill=guide_legend(title="Genera", ncol = 1, reverse = FALSE)) + theme(axis.title.x = element_text(vjust=-1.2), legend.text = element_text(face="italic"), panel.background = element_blank(), axis.line = element_line(colour = "grey"), legend.key.size = unit(0.4, "cm")) + scale_y_continuous(breaks=brks, labels = y_labels) + theme(text = element_text(size = 20))

#ggsave("stackedbarchartbouts03102023.png", plot=genusbouts, height = 8, width = 11, unit="in", dpi=500)

genusbouts
```

# Supplemental Figure 3

```{r, error=TRUE, warning=FALSE}

#PLOT ONE
plt1a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = Observed,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt1a <- plt1a + 
  # Add labels and title
  labs(
    x = "Breastfeeding bouts",
    y = "HMM Observed Richness",
    title = "HMM Richness by Breastfeeding Bouts Groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt1a

#PLOT TWO
plt1b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = Observed,
  p.adjust.method = "none",
    var.equal=TRUE, 
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt1b <- plt1b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "HMM Observed Richness",
    title = "HMM Richness by Total Time Breastfeeding Groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt1b

#PLOT THREE
plt2a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = Shannon,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt2a <- plt2a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "HMM Shannon Diversity",
    title = "HMM Shannon Diversity by Breastfeeding Bouts Groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt2a

#PLOT FOUR
  plt2b <- ggbetweenstats(
    data = md,
    x = zmomnursecutw,
    y = Shannon,
    p.adjust.method = "none",
    var.equal=TRUE,
    point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
  )
  
  plt2b <- plt2b + 
    # Add labels and title
    labs(
      x = "Total Time Breastfeeding",
      y = "HMM Shannon Diversity",
      title = "HMM Shannon Diversity by Total Time Breastfeeding Groups"
    ) + 
    # Customizations
    theme(
      # This is the new default font in the plot
      text = element_text(size = 8, color = "black"),
      plot.title = element_text(
        size = 16,
        face = "bold",
        color = "#2a475e"
      ),
      # Statistical annotations below the main title
      plot.subtitle = element_text(
        size = 15, 
        face = "bold",
        color="#1b2838"
      ),
      plot.title.position = "plot", # slightly different from default
      axis.text = element_text(size = 10, color = "black"),
      axis.title = element_text(size = 12)
    )
plt2b

#PLOT FIVE
plt3a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = shannon_evenness,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt3a <- plt3a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "HMM Shannon Evenness",
    title = "HMM Shannon Evenness by Breastfeeding Bouts Groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt3a

#PLOT SIX
plt3b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = shannon_evenness,
  p.adjust.method = "none",
    var.equal=TRUE
)

plt3b <- plt3b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "HMM Shannon Evenness",
    title = "HMM Shannon Evenness by Total Time Breastfeeding Groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt3b

#PLOT 1 for work outside home
pltw1 <- ggbetweenstats(
  data = md,
  x = workoutsidehome,
  y = Observed,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

pltw1 <- pltw1 + 
  # Add labels and title
  labs(
    x = "Maternal Work Outside Home",
    y = "HMM Observed Richness",
    title = "HMM Richness by Maternal Work Outside Home"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
pltw1

#PLOT 2 for work outside home
pltw2 <- ggbetweenstats(
  data = md,
  x = workoutsidehome,
  y = Shannon,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

pltw2 <- pltw2 + 
  # Add labels and title
  labs(
    x = "Maternal Work Outside Home",
    y = "HMM Shannon Diversity",
    title = "HMM Shannon Diversity by Maternal Work Outside Home"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
pltw2

#PLOT 3 for work outside home
pltw3 <- ggbetweenstats(
  data = md,
  x = workoutsidehome,
  y = shannon_evenness,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

pltw3 <- pltw3 + 
  # Add labels and title
  labs(
    x = "Maternal Work Outside Home",
    y = "HMM Shannon Evenness",
    title = "HMM Shannon Evenness by Maternal Work Outside Home"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
pltw3


#PLOT 7
plt4a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = zf_allophyscon,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt4a <- plt4a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "Allomother physical contact frequency",
    title = "Allomother physical contact by breastfeeding bouts groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt4a

#PLOT 8
plt4b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = zf_allophyscon,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt4b <- plt4b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "Allomother physical contact frequency",
    title = "Allomother physical contact by total time breastfeeding groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt4b

#PLOT 9
plt5a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = infecorichnessnh,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt5a <- plt5a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "Non-household caregiving network",
    title = "Non-household caregiving network by breastfeeding bouts groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt5a

#PLOT 10
plt5b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = infecorichnessnh,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt5b <- plt5b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "Non-household caregiving network",
    title = "Non-household caregiving network by total time breastfeeding groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt5b

#PLOT 11
plt6a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = agemos,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt6a <- plt6a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "Infant age (months)",
    title = "Infant age by breastfeeding bouts groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt6a

#PLOT 12
plt6b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = agemos,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt6b <- plt6b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "Infant ages (months)",
    title = "Infant age by total time breastfeeding groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt6b

#PLOT 13
plt7a <- ggbetweenstats(
  data = md,
  x = zbouts30cut,
  y = momage2,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt7a <- plt7a + 
  # Add labels and title
  labs(
    x = "Breastfeeding Bouts",
    y = "Maternal age (years)",
    title = "Maternal age by breastfeeding bouts groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt7a

#PLOT 14
plt7b <- ggbetweenstats(
  data = md,
  x = zmomnursecutw,
  y = momage2,
  p.adjust.method = "none",
    var.equal=TRUE,
  point.args=list(position = ggplot2::position_jitterdodge(dodge.width = 0.6), alpha=0.75)
)

plt7b <- plt7b + 
  # Add labels and title
  labs(
    x = "Total Time Breastfeeding",
    y = "Maternal age (years)",
    title = "Maternal age by total time breastfeeding groups"
  ) + 
  # Customizations
  theme(
    # This is the new default font in the plot
    text = element_text(size = 8, color = "black"),
    plot.title = element_text(
      size = 16,
      face = "bold",
      color = "#2a475e"
    ),
    # Statistical annotations below the main title
    plot.subtitle = element_text(
      size = 15, 
      face = "bold",
      color="#1b2838"
    ),
    plot.title.position = "plot", # slightly different from default
    axis.text = element_text(size = 10, color = "black"),
    axis.title = element_text(size = 12)
  )
plt7b

patchwork1 <- (plt1a | plt1b) / (plt2a | plt2b)  / (plt3a | plt3b) 

patchwork2 <- (plt4a | plt4b) / (plt5a | plt5b) / (plt6a | plt6b) / (plt7a | plt7b)

patchwork3 <- (pltw1 | pltw2) / (plot_spacer() + pltw3 + plot_spacer())


#ggsave(patchwork1, file="Boxplots02152023a.png", height = 14, width = 14, dpi=300)
#ggsave(patchwork2, file="Boxplots02152023b.png",  height = 20, width = 15, dpi=300)
#ggsave(patchwork3, file="Boxplots02152023c.png",  height = 20, width = 14, dpi=300)
```

